### 2015-04-28 datapaper_plants.tex
xm's first edits

\section{Data Collection}\label{sec:2}

\subsection{Plants and Growth Conditions}

{\it Arabidopsis thaliana} (ecotype Col-0) plants were grown at $20^{\circ}C$, under a $16$ hr:$8$ hr day night cycle with a daylight intensity set at $100~\mu mol~photons~m^{-2} s^{-1}$.
%
Black bean plants ({\it Phaseolus vulgaris L.}) of the cultivar Jaguar, were grown under a $14$ hr:$10$ hr day night cycle with night and day temperatures of $18^{\circ}C$ and $24^{\circ}C$ respectively, and a daylight intensity set at $200~\mu mol~photons~m^{-2} s^{-1}$.
%
Note that the bean plants were watered with half-strength Hoaglands solution three times per week.

In all cases, seeds were planted in soil covered with a black foam mask in order to minimize the fluorescence background from algal growth.
%
Two-week-old plants ({\it Arabidopsis} or bean) were transferred to imaging chambers and allowed to acclimate for $24$ hours to the LED lighting before the start of the data collection. 
Growth conditions as described above were maintained for each set of plants for the duration of image collection.