### 2015-04-29 datapaper_conclusion.tex
xm's first edits

\section{Conclusions}

This paper presents a newly collected multi-modality plant imagery database, ``MSU-PID''.
Compared to existing databases in the field, MSU-PID not only has multiple calibrated modalities, but also enables a wide variety of plant image analysis applications.
Therefore, we believe this new database will be benefitial to the research community in terms of algorithm development, performance evaluation, and identifying new research problems in plant image analysis.
Furthmore, we are also open to suggestions and comments from the users of this database to further enhance our imaging setup and capturing protocol, so that we can develop new databases in the future.