### 2015-05-02 datapaper.tex

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{bm}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{color}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[misc,geometry]{ifsym}
\usepackage{threeparttable}

%\usepackage{paralist}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}\sloppy

\title{Multi-modality Imagery Database for Plant Phenotyping \thanks{$^\star$ denotes equal contribution by the authors. This research was supported by Chemical Sciences, Geosciences and Biosciences Division, Office of Basic Energy Sciences, Office of Science, U.S. Department of Energy (award number DE-FG02-91ER20021), and by Center for Advanced Algal and Plant Phenotyping (CAAPP), Michigan State University.
}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Jeffrey A Cruz$^\star$ \and Xi Yin\footnote{Equal contributor}  \and Xiaoming Liu \and Saif M Imran \and Daniel~D~Morris \and David M Kramer    \and Jin Chen}

\authorrunning{Cruz {\it et al}} % if too long for running head

\institute{Jeffrey A Cruz $\cdot$ Jin Chen $\cdot$ David M Kramer \at
              Department of Energy Plant Research Laboratory, Michigan State University, East Lansing, MI 48824, USA  \\
		E-mail: \{cruzj,jinchen,kramerd8\}@msu.edu		
%             \emph{Present address:} of F. Author  %  if needed
           \and
           Xi Yin $\cdot$ Xiaoming Liu (\Letter)\at
              Department of Computer Science and Engineering, Michigan State University,  East Lansing, MI 48824, USA  \\
		E-mail: \{yinxi1,liuxm\}@cse.msu.edu
                      %  \\
              \and
           Saif Imran $\cdot$ Daniel D Morris \at
              Department of Electrical and Computer Engineering, Michigan State University,
              East Lansing, MI 48824, USA\\
		E-mail: \{imransai,dmorris\}@msu.edu
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}

Among many applications of machine vision, plant image analysis has recently begin to gain attention due to its potential impact on plant visual phenotyping, particularly in understanding plant growth, assessing the quality/performance of crop plants and improving crop yield. Despite its importance, the lack of publicly available research databases containing plant imagery has substantially hindered the advancement of plant image analysis. To alleviate this issue, this paper presents a new multi-modality plant imagery database named ``MSU-PID'', with two distinct properties. First, MSU-PID is captured using four types of imaging sensors, Fluorescence, infrared (IR), RGB color, and depth. Second, the imaging setup and the variety of manual labels allow MSU-PID to be suitable for a diverse set of plant image analysis applications, such as leaf segmentation, leaf counting, leaf alignment, and leaf tracking. We provide detailed information on the plants, imaging sensors, calibration, labeling, and baseline performances of this new database.
%
%Among many applications of machine vision, plant image analysis has recently begin to gain attentions due to its potential huge impact to plant visual phenotyping, the understanding of plant growth, the quality of crops, and the improvement of the yield.
%Despite its importance, the pack of publicly available research database of plant imagery has substantially hindered the advancement of plant image analysis.
%To alleviate this issue, this paper presents a new multi-modality plant imagery database named ``MSU-PID'', with two distinct properties.
%First, MSU-PID is captured using four types of imaging sensors, Fluorescence, infrared (IR), RGB color, and depth.
%Second, the imaging setup and the variety of manual labels allow MSU-PID to be suitable for a diverse set of plant image analysis applications, such as leaf segmentation, leaf counting, leaf alignment, and leaf tracking.
%We provide detailed information on the plants, imaging sensors, calibration, labeling, and baseline performances of this new database.

\keywords{Plant Phenotyping  \and Computer Vision \and Plant image \and Leaf segmentation \and Leaf tracking\and Multiple sensors \and Arabidopsis \and Bean}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\input{datapaper_intro.tex}

\input{datapaper_prior.tex}

\input{datapaper_plants.tex}

\input{datapaper_hardware.tex}

\input{datapaper_calibration.tex}

\input{datapapr_labeling.tex}

\input{datapaper_baseline.tex}

\input{datapaper_conclusion.tex}

%\paragraph{Paragraph headings} Use paragraph headings as needed.
%\begin{equation}
%a^2+b^2=c^2
%\end{equation}
%
%% For one-column wide figures use
%\begin{figure}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:1}       % Give a unique label
%\end{figure}
%%
%% For two-column wide figures use
%\begin{figure*}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics[width=0.75\textwidth]{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:2}       % Give a unique label
%\end{figure*}
%%
%% For tables use
%\begin{table}
%% table caption is above the table
%\caption{Please write your table caption here}
%\label{tab:1}       % Give a unique label
%% For LaTeX tables use
%\begin{tabular}{lll}
%\hline\noalign{\smallskip}
%first & second & third  \\
%\noalign{\smallskip}\hline\noalign{\smallskip}
%number & number & number \\
%number & number & number \\
%\noalign{\smallskip}\hline
%\end{tabular}
%\end{table}

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

{%\small
\bibliographystyle{springer}
%\bibliography{/Sharefolder/Dropbox/Bibliographies/abbrev_brief,/Sharefolder/Dropbox/Bibliographies/xl-literature}
\bibliography{abbrev,xl-literature,science}
}
\end{document}
% end of file template.tex