### 2015-07-26 datapaper_plants.tex

\section{Data Collection}
\label{sec:2}

\subsection{Plants and Growth Conditions}

{\it Arabidopsis thaliana} (ecotype Col-$0$) plants were grown at $20^{\circ}C$, under a $16~hr$ : $8~hr$ day-night cycle with a daylight intensity set at $100~\mu mol~photons~m^{-2} s^{-1}$.
%
Black bean plants ({\it Phaseolus vulgaris L.}) of the cultivar Jaguar, were grown under a $14~hr$ : $10~hr$ day-night cycle with day and night temperatures of $24^{\circ}C$ and $18^{\circ}C$ respectively, and a daylight intensity set at $200~\mu mol~photons~m^{-2} s^{-1}$.
%
Note that the bean plants were watered with half-strength Hoaglands solution three times per week.

In all cases, seeds were planted in soil covered with a black foam mask in order to minimize the fluorescence background from algal growth.
%
Two-week-old plants ({\it Arabidopsis} and bean) were transferred to imaging chambers and allowed to acclimate (\todo{Jeff: Please define ``acclimate" as in Comment $37$}) for $24$ hours to the LED lighting before the start of the data collection.
Growth conditions as described above were maintained for each set of plants for the duration of image collection.